---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# validationserver

This repository contains the R package to let users submit analyses to the Urban Institute's [Safe Data Technologies](https://www.urban.org/projects/safe-data-technologies) Validation Server Version 2.0 prototype. See the [technical white paper](https://www.urban.org/research/publication/privacy-preserving-validation-server-version-2) for an overview of the project and detailed information about the prototype. 

## Usage

### Installation

To install `validationserver` from GitHub: 

```{r gh-installation, eval=FALSE}
devtools::install_github('UrbanInstitute/validation-server-v2-r-package')
```


A valid R script must: 

1. Include a main function called `run_analysis()` that takes a `conf_data` argument with the confidential data. 
2. Use the `get_table_output()` and/or `get_model_output()` functions to define analyses. 
3. Use the `submit_output()` function to specify all analyses to run. 

Otherwise, the R script can include arbitrary code to process and analyze the underlying data. The full R script will be executed, so you can define helper functions outside of the main `run_analysis()` function. 

### Basic Example

This is a basic example of a script with a linear regression and summary table using the `penguins` dataset from the [`palmerpenguins` R package](https://allisonhorst.github.io/palmerpenguins/): 

```{r basic-example, eval=FALSE}
library(dplyr)
library(palmerpenguins)

run_analysis <- function(conf_data) {
    # Arbitrary code -----------------------------------------------------------
    transformed_data <- conf_data %>%
        filter(year == 2007) %>%
        mutate(bill_ratio = bill_length_mm / bill_depth_mm)
    
    # Specify analyses -----------------------------------------------------------
    # Example regression 
    example_fit <- lm(body_mass_g ~ bill_ratio, data = transformed_data)
    example_model <- get_model_output(
        fit = example_fit, 
        model_name = "Example Model"
    )
    
    # Example table 
    example_table <- get_table_output(
        data = transformed_data,
        table_name = "Example Table",
        stat = c("mean", "sd"),
        var = "bill_ratio"
    )
    
    # Submit analyses ------------------------------------------------------------
    submit_output(example_table, example_model)
}
```

### Model Examples 

The `get_model_output()` function requires two arguments: 

* `model_name` is an arbitrary name for the analysis. This name will show up in your output and should be unique to each analysis in an R script. 
* `fit` is a fit model object. Only `lm` and `glm` objects are currently supported.

Here's an example of a script with a linear model (fit using `lm`) and binomial model (fit using `glm`): 

```{r model-example, eval=FALSE}
library(dplyr)
library(palmerpenguins)

run_analysis <- function(conf_data) {
    # Arbitrary code -----------------------------------------------------------
    transformed_data <- conf_data %>%
        mutate(mass_above_4kg = case_when(body_mass_g > 4000 ~ 1, 
                                          TRUE ~ 0))

    # Specify analyses -----------------------------------------------------------
    # Example linear model 
    lm_fit <- lm(body_mass_g ~ bill_length_mm, data = transformed_data)
    lm_example <- get_model_output(
        fit = lm_fit, 
        model_name = "Example Linear Model"
    )

    # Example binomial model 
    glm_fit <- glm(mass_above_4kg ~ bill_length_mm, family = binomial, data = transformed_data)
    glm_example <- get_model_output(
        fit = glm_fit, 
        model_name = "Example Binomial Model"
    )

    # Submit analyses ------------------------------------------------------------
    submit_output(lm_example, glm_example)
}
```

### Table Examples 

The `get_table_output()` function includes the following arguments: 

* `data` is the data frame object to create the summary table from. To use the full data, specify `data = conf_data`. 
* `table_name` is an arbitrary name for the analysis. This name will show up in your output and should be unique to each analysis in an R script. 
* `stat` is the summary statistic (or set of summary statistics) to calculate. This can include `mean`, `median`, `sum`, `sd`, `var`, `min`, `max`, `n`, `n_distinct`, or `quantile(numeric_vector)`, where `numeric_vector` can be any numeric vector with numbers between 0 and 1 of length greater than 0; e.g. `quantile(c(0.1, 0.5, .9))` 
* `var` is the variable (or set of variables) to summarize. 
* `by` (optional) is the variable (or set of variables) to group by. 
* `na.rm` (optional) specifies explicit NA behavior (defaults to `FALSE`). 

Here's an example of a script with (1) a single summary statistic, and (2) a table demonstrating how to pass multiple values into the `stat`, `var`, and `by` arguments. 

```{r table-example, eval=FALSE}
library(dplyr)
library(palmerpenguins)

run_analysis <- function(conf_data) {
    # Arbitrary code -----------------------------------------------------------
    transformed_data <- conf_data %>%
        filter(year == 2007) %>%
        mutate(bill_ratio = bill_length_mm / bill_depth_mm)
    
    # Specify output -----------------------------------------------------------
    # Example summary statistic 
    table1 <- get_table_output(
        data = transformed_data, 
        table_name = "Example Table 1",
        stat = "mean", 
        var = "bill_ratio",
        na.rm = TRUE
    )
    
    # Example table with multiple stat/var/by values  
    table2 <- get_table_output(
        data = transformed_data, 
        table_name = "Example Table 2",
        stat = c("mean", "median", "sd", "quantile(seq(0, 1, .1))", "n"),
        var = c("bill_length_mm", "bill_ratio"),
        by = c("species", "island"), 
        na.rm = TRUE
    )
    
    # Submit output ------------------------------------------------------------
    submit_output(table1, table2)
}
```

## Contact
This work is developed by the Urban Institute. For questions, reach out to: validationserver@urban.org. 
